Contextuality and Nonlocality in Discrete Quantum Theory

### 2016-01-27 DQC3.tex
"6.1 Gleason's Theorem in DQT" Updated...

Let $\C^{*}=\C\backslash\left\{ 0\right\} $ and $\C^{d*}=\left\{ \ket{\Psi}\in\C^{d}\mid\ip{\Psi}{\Psi}\ne0\right\} $
be the set of non-zero vectors. Because normalization and global phase
gives the same state, and ${\cal P}_{\rho}:\C^{d*}\rightarrow\left[0,1\right]$
should be a state-function, we should have 
\begin{equation}
\textrm{For any vector }\ket{i}\in\C^{d*}\textrm{, and }\gamma\in\C^{*}\textrm{ , }{\cal P}_{\rho}(\ket{i})={\cal P}_{\rho}(\gamma\ket{i})\textrm{ .}\label{eq:parallel-DQT(I)}
\end{equation}
More importantly, ${\cal P}_{\rho}$ defines a probability which means
\begin{equation}
\textrm{For any orthonormal basis }\{\ket{i}\}\textrm{ , }{\displaystyle \sum_{i=0}^{d-1}{\cal P}_{\rho}\left(\ket{i}\right)=1}\textrm{ .}\label{eq:total-probability}
\end{equation}
Conversely, if a state-function ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$
is a probability, i.e., satisfying equation (\ref{eq:parallel-DQT(I)})
and (\ref{eq:total-probability}), we can identify ${\cal P}$ to
a mixed state $\rho$ by Gleason's theorem.
\begin{thm}
(Gleason's Theorem) 
\begin{enumerate}
\item For $d\ge3$, for any state-function ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$,
if ${\cal P}$ is a probability, there is a density matrix $\rho$
such that ${\cal P}={\cal P}_{\rho}$~\cite{Redhead1987-REDINA,gleason1957}.\footnote{In order to simplify the statement, we consider consider the probability
on the state vectors instead of the projectors or closed subspaces
as in the Gleason's original paper.}
\item Consequently, the Born rule (\ref{prob1.eq-6}) is the only reasonable
probability assignment, and can be deduced by other reasonable axioms~\cite{peres1995quantum,inun.606849520040101}. 
\end{enumerate}
\end{thm}
An important meaning of Gleason's theorem is that the state space
is sufficient enough to produce every possible probability we get
from an experiment. Hence, we don't have to worry about whether there
is a chance that the probability got from a measurement cannot be
explained by a state, and we need to introduce a new state. This may
be analog to the fact that every algebraic equation over $\C$ can
be solved in $\C$ so that we don't have to introduce new numbers
to solve an equation. The other important meaning of Gleason's theorem
is that it implies the Born rule (\ref{prob1.eq-6}). This is even
more important for building a discrete quantum theory. If we plug
in vectors over $\ff{p^{2}}$ into the usual Born rule (\ref{prob1.eq-6}),
it gives an elements in $\ff{p}$ which has no way interpreted as
a probability. Everyone who tried to build a quantum theory over finite
fields struggled on how to define a reasonable probability. If Gleason's
theorem works in any sense over finite fields, we can have a reasonable
way to induce the Born rule over finite fields.

>>>>

Let $\C^{*}=\C\backslash\left\{ 0\right\} $ and $\C^{d*}=\left\{ \ket{\Psi}\in\C^{d}\mid\ip{\Psi}{\Psi}\ne0\right\} $
be the set of non-zero vectors. Because normalization and global phase
gives the same state, and ${\cal P}_{\rho}:\C^{d*}\rightarrow\left[0,1\right]$
should be a state-function, we should have 
\begin{equation}
\textrm{For any vector }\ket{i}\in\C^{d*}\textrm{, and }\gamma\in\C^{*}\textrm{ , }{\cal P}_{\rho}(\ket{i})={\cal P}_{\rho}(\gamma\ket{i})\textrm{ .}\label{eq:parallel-DQT(I)}
\end{equation}
More importantly, ${\cal P}_{\rho}$ defines a probability which means
\begin{equation}
\textrm{For any orthonormal basis }\{\ket{i}\}\textrm{ , }{\displaystyle \sum_{i=0}^{d-1}{\cal P}_{\rho}\left(\ket{i}\right)=1}\textrm{ .}\label{eq:total-probability}
\end{equation}
Conversely, if a state-function ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$
is a probability, i.e., satisfying equation (\ref{eq:parallel-DQT(I)})
and (\ref{eq:total-probability}), we can identify ${\cal P}$ to
a mixed state $\rho$ by Gleason's theorem.
\begin{thm}
(Gleason's Theorem) 
\begin{enumerate}
\item For $d\ge3$, for any state-function ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$,
if ${\cal P}$ is a probability, there is a density matrix $\rho$
such that ${\cal P}={\cal P}_{\rho}$~\cite{Redhead1987-REDINA,gleason1957}.\footnote{In order to simplify the statement, we consider consider the probability
on the state vectors instead of the projectors or closed subspaces
as in the Gleason's original paper.}
\item Consequently, the Born rule (\ref{prob1.eq-6}) is the only reasonable
probability assignment, and can be deduced by other reasonable axioms~\cite{peres1995quantum}. 
\end{enumerate}
\end{thm}
An important meaning of Gleason's theorem is that the state space
is sufficient enough to produce every possible probability we get
from an experiment. Hence, we don't have to worry about whether there
is a chance that the probability got from a measurement cannot be
explained by a state, and we need to introduce a new state. This may
be analog to the fact that every algebraic equation over $\C$ can
be solved in $\C$ so that we don't have to introduce new numbers
to solve an equation. The other important meaning of Gleason's theorem
is that it implies the Born rule (\ref{prob1.eq-6}). This is even
more important for building a discrete quantum theory. If we plug
in vectors over $\ff{p^{2}}$ into the usual Born rule (\ref{prob1.eq-6}),
it gives an elements in $\ff{p}$ which has no way interpreted as
a probability. Hence, everyone who tried to build a quantum theory
over finite fields struggled on how to define a reasonable probability.
If Gleason's theorem works in any sense over finite fields, we can
have a reasonable way to induce the Born rule over finite fields.
